---
title: "Class 15: Genomics I"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Read sample genotypes data from ENSEMBLE

We downloaded genotype data from ENSEMBLE on the MXL Mexican Ancestry in Los Angeles, California dataset.

What proportion of this data are G/G etc?

```{r}
mxl <- read.csv("373531-SampleGenotypes-Homo_sapiens_Variation_Sample_rs8067378.csv")

head(mxl)
```

We want to look at the second column that contains the genotype information

```{r}
table( mxl$Genotype..forward.strand. )
```


## RNA-Seq result analysis for different genotypes of this SNP

```{r}
expr <- read.table("rs8067378_ENSG00000172057.6.txt")
head(expr)
```

```{r}
summary( expr[ expr$geno == "G/G", ]$exp )
```

```{r}
summary( expr[ expr$geno == "A/G", ]$exp )
```

```{r}
summary( expr[ expr$geno == "A/A", ]$exp )
```

## Try a boxplot

We will use the `boxplot()` function and the input data will be **expr**. How do we draw a useful plot?

```{r}
boxplot(exp ~ geno , data=expr, notch=TRUE)
```

How many samples are we looking at here?
```{r}
nrow(expr)
```


```{r}
hist( expr[ expr$geno == "G/G", ]$exp, breaks=20 )
```



```{r}
x <- 1:10
x[ x > 5 ]
```



## Try dplyr approch
```{r}
library(dplyr)
```

```{r}
expr %>%
  group_by(geno) %>%
  summarise(count_by_genotype=n(),
            mean_exp = mean(exp),
            sd_exp = sd(exp))
```

And a ggplot

```{r}
library(ggplot2)

ggplot(expr, aes(geno, exp, fill=geno)) +
  geom_boxplot(notch = TRUE, outlier.alpha=0.5) +
  labs(x="Genotype", y="Expression",
       title="Expresion of by genotype for rs8067378",
       subtitle = "See class15 workbook",
       caption="ENSG00000172057") +
  theme(legend.position = "none") 
```

And a formated table
```{r}
expr %>%
  group_by(geno) %>%
  summarise(number_samples=n(),
            mean_exp = round(mean(exp),2),
            sd_exp = round(sd(exp),2)) %>%
  kableExtra::kable()
```



