---
title: 'Class 9: Machine Learning pt1.'
author: "Barry Grant"
date: "2/5/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## K-means clustering

Let's try the `kmeans()` function in R to cluster some made-up example data.

```{r}
tmp <- c(rnorm(30,-3), rnorm(30,3))
x <- cbind(x=tmp, y=rev(tmp))

plot(x)

```

Use the kmeans() function setting k to 2 and nstart=20 
```{r}
km <- kmeans(x, centers = 2, nstart = 20)
```

Inspect/print the results 
```{r}
km
```

What is in the output object `km` I can use the `attributes()` function to find this info :-)

```{r}
attributes(km)
```

Q. How many points are in each cluster?
```{r}
km$size
```

Q. What ‘component’ of your result object details 
      - cluster size?
      - cluster assignment/membership? 
      - cluster center?

```{r}
km$cluster
```

let's check how many 2s and 1s are in this vector with the `table()` function.

```{r}
table(km$cluster)
```


Plot x colored by the kmeans cluster assignment and 
      add cluster centers as blue points
      
```{r}
plot(x, col=km$cluster)
points(km$centers, col="blue", pch=15, cex=3)
```
      

## Hierarchical clustering in R 

The `hclust()` function is the main Hierarchical clustering method in R and it **must** be passed a *distance matrix* as input, not your raw data!

```{r}
hc <- hclust( dist(x) )
hc
```

```{r}
plot(hc)
abline(h=6, col="red", lty=2)
abline(h=3.5, col="blue", lty=2)

```

```{r}
grps <- cutree(hc, h=6)
grps
```

```{r}
table( cutree(hc, h=3.5) )
```

You can also ask `cutree()` for the `k` number of groups that yopu want.

```{r}
cutree(hc, k=5)
```



```{r}
x.grp1 <- x[grps==1]
```


### Some more messy data to cluster

```{r}
# Step 1. Generate some example data for clustering
x <- rbind(
  matrix(rnorm(100, mean=0,  sd=0.3), ncol = 2),  # c1
  matrix(rnorm(100, mean=1,  sd=0.3), ncol = 2),  # c2
  matrix(c(rnorm(50, mean=1, sd=0.3),             # c3
           rnorm(50, mean=0, sd=0.3)), ncol = 2))
colnames(x) <- c("x", "y")

# Step 2. Plot the data without clustering
plot(x)

# Step 3. Generate colors for known clusters 
#         (just so we can compare to hclust results)
col <- as.factor( rep(c("c1","c2","c3"), each=50) ) 

plot(x, col=col)
```

Q. Use the dist(), hclust(), plot() and cutree()
      functions to return 2 and 3 clusters 

```{r}
hc <- hclust( dist(x) )
plot(hc)
```
  
```{r}
grps3 <- cutree(hc, k=3)
grps3
table(grps3)
```
    
Q. How does this compare to your known 'col' groups?

```{r}
plot(x, col=grps3)
```

```{r}
table(grps3, col)
```


# Principal Component Analysis (PCA)

The main function in base R for PCA is called `prcomp()`. Here we will use PCA to examine the funny food that folks eat in the UK and N. Ireland.

Import the CSV file first:

```{r}
x <- read.csv("UK_foods.csv", row.names = 1)
x
```

Make some conventional plots
```{r}
barplot(as.matrix(x), beside=F, col=rainbow(nrow(x)))
```


```{r}
barplot(as.matrix(x), beside=T, col=rainbow(nrow(x)))
```


```{r}
pairs(x, col=rainbow(10), pch=16)
```

# PCA to the rescue!

```{r}
pca <- prcomp( t(x) )
```

```{r}
summary(pca)
```

```{r}
attributes(pca)
```


```{r}
plot( pca$x[,1], pca$x[,2], xlab="PC1 (67.4%)", ylab="PC2 (29%)" )
text(pca$x[,1], pca$x[,2], labels = colnames(x), 
     col=c("black","red", "blue", "darkgreen"))
```


```{r}
plot(pca)
```

