---
title: "Structural Bioinformatics (pt. 1)"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## PDB Statistics

Download a CSV file from the PDB site (accessible from “Analyze” > “PDB Statistics” > “by Experimental Method and Molecular Type”. Move this CSV file into your RStudio project and use it to answer the following questions:

> Q: What percentage of structures in the PDB are solved by X-Ray and Electron Microscopy.

```{r}
stats <- read.csv("Data Export Summary.csv", row.names = 1)
stats
```

```{r}
ans <- round(stats$Total / sum(stats$Total) * 100, 2)
names(ans) <- rownames(stats)
ans
```


> Q: What proportion of structures in the PDB are protein?

```{r}
round(sum(stats$Proteins) / sum(stats$Total) * 100, 2)
```


```{r}

```

> Q: Type HIV in the PDB website search box on the home page and determine how many HIV-1 protease structures are in the current PDB?


## 2. Using bio3d to analyse protein structure data

Use the bio3d package from CRAN. 

```{r}
library(bio3d)

pdb <- read.pdb("1hsg")
pdb
```

Adenylate kinase structures from the PDB for analysis

```{r}
hits <- NULL
hits$pdb.id <- c('1AKE_A','4X8M_A','6S36_A','6RZE_A','4X8H_A',
                 '3HPR_A','1E4V_A','5EJE_A','1E4Y_A','3X2S_A',
                 '6HAP_A','6HAM_A','4K46_A','4NP6_A','3GMT_A',
                 '4PZL_A')
```

Download and parse these structures

```{r}
# Download releated PDB files
files <- get.pdb(hits$pdb.id, path="pdbs", split=TRUE, gzip=TRUE)
```


Align and superpose structures

```{r}
# Align releated PDBs
pdbs <- pdbaln(files, fit = TRUE, exefile="msa")
```



```{r}
library(rgl)
library(bio3d.view)
view(pdbs)

rglwidget(elementId = "tmp")
```



```{r}
pc <- pca(pdbs)
plot(pc)
```



```{r}
plot(pc, 1:2)
```


Make a PCA trajectory 


```{r}
pc1 <- mktrj(pc, pc=1, file="pc_1.pdb")

```



```{r}
view(pc1)
```

