---
title: "Structure based drug design"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Download and process starting structure

Here we download and clean up the HIV-Pr structure (PDB code: 1HSG)
from the main PDB database. We will make a seperate set of "protein-only" and "ligand only" PDB files.

```{r}
library(bio3d)
file.name <- get.pdb("1hsg")
```

We will use `read.pdb()`, `atom.select()` and `write.pdb()` functions to make our seperate "protein-only" and "ligand only" PDB files

```{r}
hiv <- read.pdb(file.name)
hiv
```

```{r}
prot <- atom.select(hiv, "protein", value=TRUE)
write.pdb(prot, file="1hsg_protein.pdb")
```

And ligand only file 

```{r}
lig <- atom.select(hiv, "ligand", value = TRUE)
lig
write.pdb(lig, file="1hsg_ligand.pdb")
```

## Read docking results

Read in the output of docking and make a PDB file for viewing in VMD or PyMol.

```{r}
res <- read.pdb("all.pdbqt", multi=TRUE) 
write.pdb(res, "results.pdb")
```

























